\documentclass[letterpaper,11pt]{article}
%\documentclass[12pt]{article}
%\usepackage[margin=0.5in]{geometry}
%\usepackage{hyperref}
%\usepackage{tikz}
%\usepackage{amsmath}
%\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
%\newcommand{\unit}[1]{\ensuremath{\, \mathrm{#1}}}
%\newcommand{\inlinecode}{\texttt}

\begin{document}

\title{CS 830: Program 4}
\date{April 10, 2013}
\author{Carmen St.\ Jean}

\maketitle

\begin{enumerate}
\item \emph{Describe any implementation choices you made that you felt were important. Mention anything else that we should know when evaluating your program.}

I stored most everything in 1D and 2D arrays.  I really didn't make use of much objected oriented design.  I used the exploration function wherever possible with a $k$ of 4.  I used 0.001 as the value to calculate when to stop updating $U$.
\item \emph{What can you say about the the time and space complexity of your
program?}

Let $a$ be the number of actions available and $s$ be the number of states available in a reinforcement learning problem.

Random will have a constant time complexity, since actions are selected at random.  Expected value or utility is never calculated.  Space complexity is constant as well since no information is stored and learned.

Greedy will have a space complexity of $O(s^2 * a)$, since the transition function must be kept in memory.  Its time complexity will also be $O(s * a)$ because of the calculation of the expected value.

Value iteration will have a space complexity of $O(s^2 * a)$ because each ($s$, $a$, $s'$) triplet must be represented in memory.  It's time complexity for each iteration will be $O(s^2 * a)$, though this partially depends on the discount factor.  A larger discount factor, near 1, will make decisions be based on results further and further in the future, so the number of iterations could grow exponentially.

Policy iteration will have a time complexity of $O(a)$ for action selection but $O(s^2 * a)$ for the update of PI.  Its space complexity will be $O(s^2 * a)$ because of the storing of the utility.
\item \emph{What suggestions do you have for improving this assignment in the future?}

Verification for this assignment was impossible because the reference was incomplete.  I felt ill-prepared because the slides were very unclear about variables like $k$, $loss$, $bound$, and $N$ that were not clearly defined.  Otherwise, the equations on the slides seemed incomplete to me and I had to constantly reference the book.  I did not have time to rewatch videos as we really only had a few days for this assignment compared to the others.  I don't know what I gained for this assignment actually because I am not sure if my answers are close to correct at all.
\end{enumerate}

\end{document}
